---
title: "Homework 1"
author: "Ruxin Shen"
date: "due 9/25/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE)
```

Note: Grading is based both on your graphs and verbal explanations. Follow all best practices as discussed in class.

Read *Graphical Data Analysis with R*, Ch. 3

### 1. Salary
[15 points]
```{r import}
library(dplyr)
library(tidyverse)
library(reshape2)
```

a) Draw multiple boxplots, by year, for the `Salaries` variable in *Employee.csv* (Available in the Data folder in the Files section of CourseWorks, original source: https://catalog.data.gov/dataset/employee-compensation-53987). How do the distributions differ by year?

```{r}
employees <- read.csv("Employee.csv")
year=factor(employees$Year)

g1 <- ggplot(employees, aes(x=year, y=Salaries))+
  geom_boxplot()

g1

#The spread of salaries are mostly the same, but the outliers are more 
#and more spreaded out over the years. There are more outliers on the top.
```

b) Draw histograms, faceted by year, for the same data. What additional information do the histograms provide?
```{r}
g2 <- ggplot(employees, aes(Salaries))+
  geom_histogram()+
  facet_wrap(~Year)

g2
#From the histograms, we can see the multimodality(ie, how many peaks), 
#and the distribution overviews. We can also tell the count of each bin
#so the comparison of the count of certain salary range would be possible.
```

c) Plot overlapping density curves of the same data, one curve per year, on a single set of axes. Each curve should be a different color. What additional information do you learn? 
```{r}
#melt_employees <- melt(employees)
g3 <- ggplot(employees, aes(x=Salaries, group=factor(Year), color=factor(Year)))+
            geom_density(alpha=0.25)

g3
#The additional information we learned is density. We can see clearly where 
#the peaks are and the multimodality of the distribution
```

d) Sum up the results of a), b) and c): what kinds of questions, specific to this dataset, would be best answered about the data by each of the three graphical forms?
```{r}
#Boxplot: compare the spread and see the center(median) and outliers of the data
#Histogram: compare specific count of certain salary range(ie, one bin)
#Density curve: comepare how many peaks the distribution have, and see if the
#               density of distribution changes over the years
```

### 2. Overtime

[10 points]

a) Draw multiple horizontal boxplots, grouped by `Organization Group` for the `Overtime` variable in  *Employee.csv* The boxplots should be sorted by group median.  Why aren't the boxplots particularly useful?
```{r}
g1 <- ggplot(employees, aes(x=fct_reorder(Organization.Group,Overtime,fun=median),
                          y=Overtime))+
  geom_boxplot()+
  coord_flip()

g1
#Because while the center of each group is close to the median, the outliers 
#are very spreaded out. What's more, the numbers of observations of each group
#are different. 
```

b) Either subset the data or choose another graphical form (or both) to display the distributions of `Overtime` by `Organization Group` in a more meaningful way. Explain how this form improves on the plots in part a).
```{r}
g2 <- ggplot(employees, aes(Overtime))+
  geom_histogram()+
  facet_wrap(~fct_reorder(Organization.Group,
                         Overtime,fun=median))

g2
#With histograms, we can see the differences between the number of observations
#in different groups very clearly. In addition, we can see the distribution 
#of each group more clearly see how clustered the data around median is.
```


### 3. Boundaries

[10 points]

a) Find or create a small dataset (< 100 observations) for which right open and right closed histograms for the same parameters are not identical. Display the full dataset (that is, show the numbers) and the plots of the two forms.
```{r}
a = c(1:30)
b = c(1:30)
hist(a)
hist(b, right=FALSE)
```

b) Adjust parameters--the same for both--so that the right open and right closed versions become identical. Explain your strategy.
```{r}
hist(a, breaks=seq(-2.5,32.5,5))
hist(b, right=FALSE, breaks=seq(-2.5,32.5,5))

#if we change the boundaries to half values(ie, 0.5) and all the data points
#are integers, then there would be no confusing boundary points.
```

### 4. Beavers

[10 points]

a) Use QQ (quantile-quantile) plots with theoretical normal lines to compare `temp` for the built-in *beaver1* and *beaver2* datasets.  Which appears to be more normally distributed?
```{r}
library(qqplotr)
library(datasets)

qqnorm(beaver1$temp)
qqline(beaver1$temp)

qqnorm(beaver2$temp)
qqline(beaver2$temp)

#beaver1 seems more normally distributed, althought it has fat tails, the 
#center seems fits the qqline(normal distribution). On the other side, 
#beaver2 seems more spreaded out in the center and it seems to have two 
#spikes of identical values.
```
b) Draw density histograms with density curves and theoretical normal curves overlaid. Do you get the same results as in part a)?
```{r}
g1 <- ggplot(beaver1)+
  geom_histogram(aes(x=temp,y=..density..))+
  geom_density(aes(x=temp,y=..density..))+
stat_function(fun = dnorm, args = list(mean = mean(beaver1$temp), 
                                       sd = sd(beaver1$temp)))
g1


g2 <- ggplot(beaver2)+
  geom_histogram(aes(x=temp,y=..density..))+
  geom_density(aes(x=temp,y=..density..))+
stat_function(fun = dnorm, args = list(mean = mean(beaver2$temp), 
                                       sd = sd(beaver2$temp)))
g2

#Yes, I got the same results. But density histograms and density curves are 
#more straightforward so I can see that beaver1 has really fat tails and 
#beaver2 has two peaks.
```

c) Perform the Shapiro-Wilk test for normality using the `shapiro.test()` function. How do the results compare to parts a) and b)? 
```{r}
shapiro.test(beaver1$temp)
shapiro.test(beaver2$temp)
#Since both p-values < 0.02, according to the shaprio-wilk test, the 
#distributions of the temp characteristic of the data set beaver1 and 
#beaver2 are unlikely to be normal
#The test only tell us that both data sets are unlikely normal, it doesn't 
#give us any other information on the distribution
```

### 5. Doctors

[5 points]

Draw two histograma of the number of deaths attributed to coronary artery disease among doctors in the *breslow* dataset (**boot** package). *(Hint: think carefully about the form in which you are receiving the data.)*
```{r}
library(boot)

if_smoke <- c("0"="non-smoker","1"="smoker")

ggplot(breslow,aes(x=age,y=y))+
  geom_histogram(stat="identity") + facet_grid(~smoke,
                                               labeller = labeller(smoke = if_smoke)) + ylab("deaths")
```

```{r another method}
ns <- filter(breslow,smoke==0)
s <- filter(breslow,smoke==1)
sum_nonsmoker = sum(ns$y)
sum_smoker = sum(s$y)
df <- data.frame(if_smoke=c("non_smoker","smoker"), deaths = c(sum_nonsmoker,sum_smoker))
ggplot(df, aes(x=if_smoke,y=deaths )) + 
  geom_histogram(stat="identity")
```